\documentclass[12pt]{article}
\usepackage{fancyvrb,amsmath,amsfonts,amssymb,graphicx,listings}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
%\usepackage{parskip}
\usepackage{slashed}
\usepackage{tikz}
\usepackage{mathrsfs}

% change margins
\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}
\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}

\hyphenpenalty=10000

\begin{document}

\begin{center}
{\sc starting apache}
\end{center}

\noindent
{\bf Step 1. Edit ``httpd.conf'' file and change 3 lines.}

\begin{Verbatim}
sudo vi /etc/apache2/httpd.conf

-#LoadModule socache_shmcb_module libexec/apache2/mod_socache_shmcb.so
+LoadModule socache_shmcb_module libexec/apache2/mod_socache_shmcb.so

-#LoadModule ssl_module libexec/apache2/mod_ssl.so
+LoadModule ssl_module libexec/apache2/mod_ssl.so

-#Include /private/etc/apache2/extra/httpd-ssl.conf
+Include /private/etc/apache2/extra/httpd-ssl.conf
\end{Verbatim}

\noindent
{\bf Step 2. Edit ``httpd-ssl.conf'' file and change 2 lines.}

\begin{Verbatim}
sudo vi /etc/apache2/extra/httpd-ssl.conf

-SSLCertificateFile "/private/etc/apache2/server.crt"
+SSLCertificateFile "/private/etc/apache2/my-cert.pem"

-SSLCertificateKeyFile "/private/etc/apache2/server.key"
+SSLCertificateKeyFile "/private/etc/apache2/my-key.pem"
\end{Verbatim}

\noindent
{\bf Step 3. Create a certificate and key using the following bash script.}

\begin{Verbatim}
#!/bin/bash
name=localhost
openssl genrsa -out my-key.pem 2048
openssl req -new -key my-key.pem -out my-csr.pem -subj "/O=$name/CN=$name"
echo -e "basicConstraints=critical,CA:true,pathlen:0\nsubjectAltName=DNS:$name"\
 >ext.tmp
openssl x509 -req -days 3650 -in my-csr.pem -signkey my-key.pem -out my-cert.pem\
 -extfile ext.tmp
rm ext.tmp
\end{Verbatim}

\noindent
{\bf Step 4. Copy certificate files to ``apache2'' directory.}

\begin{Verbatim}
sudo cp my-cert.pem my-key.pem /etc/apache2
\end{Verbatim}

\noindent
{\bf Step 5. Start Apache.}

\begin{Verbatim}
sudo apachectl start
\end{Verbatim}

\noindent
{\bf More info}

\noindent
Start Apache automatically on boot.

\begin{Verbatim}
sudo launchctl load -w \
 /System/Library/LaunchDaemons/org.apache.httpd.plist
\end{Verbatim}

\noindent
Stop Apache from starting on boot.

\begin{Verbatim}
sudo launchctl unload -w \
 /System/Library/LaunchDaemons/org.apache.httpd.plist
\end{Verbatim}

\noindent
Print certificate file.

\begin{Verbatim}
openssl x509 -in cert.pem -noout -text
\end{Verbatim}

\end{document}
